---
title: "Basic Data Visualization"
author: "by Kozo Nishida"
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    toc: true
    toc_depth: 1
    df_print: paged
---
```{r, echo = FALSE}
knitr::opts_chunk$set(
  eval=FALSE
)
```

# Installation
```{r, eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

if(!"RCy3" %in% installed.packages())
  BiocManager::install("RCy3")

library(RCy3)
```

## Prerequisites
In addition to the RCy3 package, you will need the latest version of Cytoscape, which can be downloaded from http://www.cytoscape.org/download.php. Simply follow the installation instructions on screen.

## Getting started
First, launch Cytoscape and keep it running whenever using RCy3. Confirm that you have everything installed and running:
```{r}
    cytoscapePing()
    cytoscapeVersionInfo()
```

# Basic Data Visualization

**Cytoscape is an open source software platform for integrating, visualizing, and analyzing measurement data in the context of networks.**

This tutorial presents a scenario of how expression and network data can be combined to tell a biological story and includes these concepts:

 - Visualizing networks using expression data.
 - Filtering networks based on expression data.
 - Assessing expression data in the context of a biological network.

# Loading Network
You can download the [demo network session file](https://nrnb.org/data/BasicDataVizDemo.cys) to your current working directory by running...
```{r}
download.file("https://nrnb.org/data/BasicDataVizDemo.cys", "./BasicDataVizDemo.cys", 'curl')
```

Now open the demo network using...
```{r}
openSession(file.location = "./BasicDataVizDemo.cys")
```

