---
title: "Basic Data Visualization"
output: html_notebook
---

# Basic Data Visualization

**Cytoscape is an open source software platform for integrating, visualizing, and analyzing measurement data in the context of networks.**

This tutorial presents a scenario of how expression and network data can be combined to tell a biological story and includes these concepts:
- Visualizing networks using expression data.
- Filtering networks based on expression data.
- Assessing expression data in the context of a biological network.

# Loading Network

- To get started,
  - install RCy3 with running

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("RCy3")
```

  - install and launch the latest version of [Cytoscape](http://www.cytoscape.org/).
- Download the [demo network session file](http://nrnb.org/data/BasicDataVizDemo.cys).
- Open the demo network using

```{r}
library(RCy3)
openSession(file.location = "./BasicDataVizDemo.cys")
```

