---
title: "Importing Data From Tables"
author: "Kozo Nishida, Kristina Hanspers and Alex Pico"
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    toc: true
    toc_depth: 4
    df_print: paged
---
```{r, echo = FALSE}
knitr::opts_chunk$set(
  eval=FALSE
)
```

*The R markdown is available from the pulldown menu for* Code *at the upper-right, choose "Download Rmd", or [download the Rmd from GitHub](https://raw.githubusercontent.com/nrnb/gsod2019_kozo_nishida/master/importing-data-from-tables.Rmd).*

<hr />

Cytoscape supports the ability to link data to networks as node and edge attributes, enabling data-driven network visualization. This tutorial describes how to load expression data from a spreadsheet, and how to link it to already loaded networks.

<hr />

# Installation
```{r, eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

if(!"RCy3" %in% installed.packages())
  BiocManager::install("RCy3")

library(RCy3)
```

# Getting started
First, launch Cytoscape and keep it running whenever using RCy3. Confirm that you have everything installed and running:
```{r}
    cytoscapePing()
    cytoscapeVersionInfo()
```

# Prerequisites
In addition to the RCy3 package, you will need the latest version of Cytoscape, which can be downloaded from http://www.cytoscape.org/download.php. Simply follow the installation instructions.

You will also need the `httr` R package to save the remote network data:

```{r}
install.packages("httr")
```

# Loading Network

- Download the network [galFiltered.sif](https://cytoscape.github.io/cytoscape-tutorials/protocols/data/galFiltered.sif).

```{r}
library(httr)
sif_url = "https://cytoscape.github.io/cytoscape-tutorials/protocols/data/galFiltered.sif"
GET(sif_url, write_disk(tf <- tempfile(fileext = ".sif")))

#df <- read_excel(tf)
```

- And load the network via

```{r}
importNetworkFromFile(file = tf)
```

# Loading Expression Data

The data you are working with is from yeast, the genes **Gal1**, **Gal4**, and **Gal80** are all yeast transcription factors.
Your expression experiments all involve some perturbation of these transcription factor genes.


Gal1, Gal4, and Gal80 are also represented in your interaction network, where they are labeled according to yeast locus tags: **Gal1 corresponds to YBR020W, Gal4 to YPL248C, and Gal80 to YML051W**.

