---
title: "Cytoscape StringApp"
author: "Kozo Nishida, Kristina Hanspers and Alex Pico"
date: "`r Sys.Date()`"
output:
  BiocStyle::html_document:
    toc: true
    toc_depth: 4
    df_print: paged
---
```{r, echo = FALSE}
knitr::opts_chunk$set(
  eval=FALSE
)
```

*The R markdown is available from the pulldown menu for* Code *at the upper-right, choose "Download Rmd", or [download the Rmd from GitHub](https://raw.githubusercontent.com/cytoscape/cytoscape-automation/master/for-scripters/R/notebooks/Cancer-networks-and-data.Rmd).*

<hr />

In these exercises, we will use the [stringApp](http://apps.cytoscape.org/apps/stringApp) for [Cytoscape](http://cytoscape.org/) to retrieve molecular networks from the [STRING](https://string-db.org/) and [STITCH](http://stitch-db.org/) databases. The exercises will teach you how to:

- retrieve networks for proteins or small-molecule compounds of interest
- retrieve networks for a disease or an arbitrary topics in PubMed
- layout and visually style the resulting networks
- import external data and map them onto a network
- perform enrichment analyses and visualize the results
- merge and compare networks
- select proteins by attributes
- identify functional modules through network clustering

The original version of this tutorial was developed by Lars Juhl Jensen of the Novo Nordisk Center for Protein Research at the University of Copenhagen. We thank professor Jensen for his gracious willingness to allow us to repackage the content for delivery as a Cytoscape tutorial.

<hr />

# Installation
```{r, eval = FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

if(!"RCy3" %in% installed.packages())
  BiocManager::install("RCy3")

library(RCy3)
```

# Getting started
First, launch Cytoscape and keep it running whenever using RCy3. Confirm that you have everything installed and running:
```{r}
    cytoscapePing()
    cytoscapeVersionInfo()
```

# Prerequisites

The exercises require you to have certain Cytoscape apps installed.

```{r}
installApp('stringApp')
installApp('yfileslayoutalgorithms')
installApp('enhancedGraphics')
installApp('clusterMaker2')
```

If you are not already familiar with the STRING database, we highly recommend that you go through the short [STRING exercises](https://jensenlab.org/training/string/) provided by the [Jensen lab](https://jensenlab.org/) to learn about the underlying data before working with them in these exercises.

